---
title: "Simple Insurance Claims Simulation"
date: "2017-03-20"
output: html_document
tut_category: "simulation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = "#>")
```

## A Single Frequency/Severity Observation

Warning: This tutorial is just a placeholder so don't working through it.

We will start running one frequency severity simulation.  We will firts simulate the frequency (i.e. the number of claims), and then we will simulate a severity for each simulated claim.

### Step 1: simulate a single frequency

We are using the poisson distriubtion to simulate the frequency.  The poisson distribution has one parameter (`lambda`).  Both the mean and the variance of the poisson distribution are equal to `lambda`.  Without further adu let's run the simulation:

```{r set_seed}
set.seed(3000)

freq <- rpois(n = 1, lambda = 10)
freq
```

We simulated `r freq` claims.

### Step 2: Severity Simulation

Now we will simulate a severity for each of these `r freq` claims.  We will use the lognormal distributions here.

```{r single_severity_simulation}
severity <- rlnorm(n = freq, meanlog = 9, sd = 1.9)
severity
```

### Step 3: Inspect the Results

```{r single_results}
summary(severity)
```

Ok. That looks pretty good.  Now let's do what computers do best and scale the number of computations up a bit.

## Multiple Frequency/Severity Observations

### Step 1: simulate the frequencies

```{r frequency_simulation}
# set the number of observations
n_obs <- 1000

# run the frequency simulations
freq <- rpois(n_obs, 10)

# inspect some of the simulated frequencies
freq[1:20]
```

### Step 2: Simulate the Severities

Now we are going to run severity simulations for each of our simulated frequencies.  This is where things may start to get a little be complicated if you are just starting out.

```{r severity_simulation}
severity <- lapply(freq, function(x) {
  rlnorm(n = x, meanlog = 9, sd = 1.9)
})
```

In the above code we used the `lapply()` function to run the `rlnorm()` function on each of the elements in out `freq` vector.  The output is a list with `r n_obs` elements.  Each element in our list output contains the number of simulated severities as specified in the `freq` vector.  e.g. the first two frequency simulations were `r freq[1:2]` and the first two elements in our `severity` list are:

```{r}
severity[1:2]
```

You can see above that `severity[[1]]` has `r freq[1]` simulated values and `severity[[2]]` has `r freq[2]` simulated values.  

## Step 3: Tidy the Data


