---
title: "Introduction to Simulations in R"
date: "2017-05-12"
output: html_document
tut_category: "simulation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = "#>")
```

The first thing that I did when learning R was run some random simulations.  It was a great way to quickly get familiar with R's base data structures and functions without needing to load data into R.  I have also found it very useful for creating sample data and coming up with quick confidence levels.  

Before you start generating random numbers, it is a good idea to set the seed so that your simulations will be reproducible.

```{r}
set.seed(1234)
```

Now that we have our seed set we can do a very basic simulation using the `sample()` function.

```{r cars}
sample(1:10, size = 2, replace = TRUE)
```

Above we drew two random observations from the numbers `r 1:10` with replacement.  It may seem pretty boring, but the `sample()` function definitely comes in handy.  Particularly for creating mock data sets.  

R also comes with a variety of probability distributions that can generate random numbers.  We will be using several of them in the upcoming tutorials.  At the end of these tutorials we will put our simualtion skills to test by creating a Shiny application that simulates and plots random numbers.
