---
title: "Simple Insurance Claims Simulation"
date: "2017-03-20"
output: html_document
tut_category: "simulation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = "#>")
```

## A Single Frequency/Severity Observation

We will start by running one frequency severity simulation.  We will first simulate the frequency (i.e. the number of claims), and then we will simulate a severity for each simulated claim.

### Step 1: simulate a single frequency

We are using the poisson distriubtion to simulate the frequency.  The poisson distribution has one parameter (`lambda`).  Both the mean and the variance of the poisson distribution are equal to `lambda`.  Without further ado let's run the simulation:

```{r set_seed}
set.seed(3000)

freq <- rpois(n = 1, lambda = 10)
freq
```

We simulated `r freq` claims.

### Step 2: Severity Simulation

Now we will simulate a severity for each of these `r freq` claims.  We will use the lognormal distribution here.

```{r single_severity_simulation}
severity <- rlnorm(n = freq, meanlog = 9, sd = 1.9)
severity
```

### Step 3: Inspect the Results

```{r single_results}
summary(severity)
```

Ok. That looks pretty good.  Now let's do what computers do best and scale the number of computations up a bit.

## Multiple Frequency/Severity Observations

### Step 1: simulate the frequencies

```{r frequency_simulation}
# set the number of observations
n_obs <- 1000

# run the frequency simulations
freq <- rpois(n_obs, 10)

# inspect some of the simulated frequencies
freq[1:20]
```

### Step 2: Simulate the Severities

Now we are going to run severity simulations for each of our simulated frequencies.  This is where things may start to get a little complicated if you are just starting out.

```{r severity_simulation}
severity <- lapply(freq, function(x) {
  rlnorm(n = x, meanlog = 9, sd = 1.9)
})
```

In the above code we used the `lapply()` function to run the `rlnorm()` function on each of the elements in out `freq` vector.  The output is a list with `r n_obs` elements.  Each element in our list output contains the number of simulated severities as specified in the `freq` vector.  e.g. the first two frequency simulations were `r freq[1:2]` and the first two elements in our `severity` list are:

```{r}
severity[1:2]
```

You can see above that `severity[[1]]` has `r freq[1]` simulated values and `severity[[2]]` has `r freq[2]` simulated values.  

## Step 3: Tidy the Data

Our data is in a list format, but we want it in a data frame.  The `dplyr` package is amazing for working with data frames.  We will be using it a lot in these tutorials.

```{r}
library(dplyr)
# first convert each list item to a data frame
severity <- lapply(1:length(severity), function(x) {
  data.frame("sim" = x, "severity" = severity[[x]])
})

# now combine the data frames into one big data frame
severity <- bind_rows(severity)
head(severity)
```

Now that we have our data in a nice tidy data frame we can do easily work with it.  First lets find the total aggregate loss for each frequency severity simulation.

```{r}
agg <- severity %>% group_by(sim) %>%
               summarize(agg = sum(severity))

head(agg)
```

The average aggregate loss is

```{r}
mean(agg$agg)
```

Now what happens if we apply a limit of 100,000 to each severity:

```{r}
agg_limited <- severity %>%
                 mutate(severity_limited = pmin(severity, 100000)) %>%
                 group_by(sim) %>%
                 summarize(agg = sum(severity),
                           agg_limited = sum(severity_limited))
head(agg_limited)
```

Our mean frequency sevrity simulation with each severity limited to 100,000 is:

```{r}
mean(agg_limited$agg_limited)
```

Here we learned the basics of a frequency/severity simulation.  In the next tutorail we will apply an aggregate stop loss, and we will introduce other distributions. 

