---
title: "About R Actuary"
author: "Andy Merlino"
---

RActuary develops open source packages and educational material to facilitate actuarial analysis in R.  RActuary is a project by [Tychobra, LLC](http://tychobra.com/).  

## R

R is a powerful platform for data analysis, and it continues to improve.  We believe R's current capabilities, its rate of improvement, and its ability to adapt to the changing data science landscape make it ideal for the future of actuarial analysis.  You can get a sense of the acceleration of innovation in R when you look at growth in packages on [CRAN](https://en.wikipedia.org/wiki/R_(programming_language)#CRAN).

```{r, echo = FALSE }
num_pkg <- read.csv("./blog/about/num-r-packages.csv", stringsAsFactors = FALSE)
num_pkg <- num_pkg[, -1]
num_pkg$date <- as.Date(num_pkg$date)

current_pckg_num <- data.frame(
  date = Sys.Date(),
  pkgs = nrow(available.packages(repos='http://cran.us.r-project.org'))
)

num_pkg <- rbind(num_pkg, current_pckg_num)
```

```{r, message = FALSE, eval = FALSE, echo = FALSE }
library(plotly)
library(dplyr)
library(lubridate)

num_pkg_2007 <- num_pkg %>%
                  mutate(year = year(date)) %>%  
                  filter(year %in% c("2007", "2017")) %>%
                  group_by(year) %>%
                  top_n(-1, date)

mark_2007 <- list(
  x = num_pkg_2007$date,
  y = num_pkg_2007$pkgs,
  text = ~paste(format(num_pkg_2007$pkgs, big.mark = ","), " Packages"),
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 7,
  ax = -20,
  ay = -40
)

g <- plot_ly(num_pkg, x = ~date, y = ~pkgs, mode = "line") %>%
       layout(title = "Packages on CRAN",
              xaxis = list(title = "Year"),
              yaxis = list(title = "Packages"),
              annotations = mark_2007) %>%
       config(displaylogo = FALSE, displayModeBar = FALSE)
       
g
#htmlwidgets::saveWidget(g, file = "./blog/about/cran-pkgs.html", selfcontained = TRUE)
```

<iframe src="../blog/about/cran-pkgs/index.html" width = "100%" height="500" ></iframe>

CRAN was created in 1997, so it took 10 years to get to 1,000 packages in 2007 and 10 more year to get to 10,000 packages in 2017.  The number of packages on CRAN is doubling every 3 years. Will CRAN have 100,000 packages in 2027?

## Actuaries

Most actuaries think R is interesting, some actuaries even use R regularly, but as of 2017 it is rare to find an actuary who uses R (or any programmed data analysis) more than he or she uses Excel.

By transitioning more of our analyses to R we can unlock the ever expanding power of the R ecosystem.  We just need to build out the tools and spread the knowledge.  Major improvments are within our grasps.

The largest and most immediate improvements to actuarial analysis can be made by using R to generate reports, communicate via the web, and work in reproducible workflows.  To help actuaries use R efficiently, we create open source R packages and projects that focus first on practical actuarial analysis (such as report creation) and second on advances in actuarial modeling.

## Rstudio

We use Rstudio's products and packages and we try to emulate the Rstudio approach to data science. If you are not familiar with Rstudio, we highly recommend the resources on their [website](https://www.rstudio.com/).

## Join Us

We are just getting started, but would love your feedback.  Please check us out on [GitHub](https://github.com/Ractuary) and let us know what you are up to: andy.merlino@tychobra.com
